---
title: "m158-project1-sara-ian"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, warning = FALSE,
                      fig.width = 5, fig.height = 3.5, 
                      fig.align = 'center')

library(tidyverse)
library(broom)
library(tidymodels)
library(mosaic)
library(NHANES)
library(nhanesA)
library(skimr)
```

## R Markdown

```{r NHANES DATA}
NHANES %>%
  glimpse()
```


We are using data from the National Health and Nutrition Exam Survey (NHANES) database that was collected between 2009-2012 with adjusted weighing.


There are 75 total variables, however only 10 are relevant to our project.

**Variables we're using**:

Age (quantitative)

Income (numerical discrete?)

Testosterone (quantitative)

Physactivedays (numerical discrete)

Pulse (quantitative)

Gender (categorical)

SleepTrouble (categorical) 

Depressed (categorical)

Education (categorical)

Work (categorical)

The target population of NHANES is "the non-institutionalized civilian resident population of the United States". Therefore, the observational unit is a civilian resident of the United States (of any age).

```{r}
OurData <- NHANES %>% select(Age, HHIncome, Testosterone, PhysActiveDays, 
                                 Pulse, Gender, SleepTrouble, Depressed, 
                                 Education, Work)
names(OurData)
dim(OurData)

```

```{r}
OurData %>%
  glimpse()
```


```{r eval = FALSE}

skim(OurData)

```

**Graphs**

```{r}
OurData %>%
  ggplot(aes(x = Pulse)) + # which column are the residuals?
  geom_histogram()  # nothing to do here, but remember this function for later!
```

```{r}
OurData %>%
  ggplot(aes(x = Testosterone)) + # which column are the residuals?
  geom_histogram()  # nothing to do here, but remember this function for later!
```

```{r}
OurData %>%
  ggplot(aes(x = PhysActiveDays)) + # which column are the residuals?
  geom_histogram()  # nothing to do here, but remember this function for later!
```

```{r}
OurData %>%
  ggplot(aes(x = Age)) + # which column are the residuals?
  geom_histogram()  # nothing to do here, but remember this function for later!
```

```{r}
OurData %>%
  ggplot(aes(x = Gender)) + # which column are the residuals?
  geom_bar()  # nothing to do here, but remember this function for later!
```

```{r}
OurData %>%
  ggplot(aes(x = SleepTrouble)) + # which column are the residuals?
  geom_bar()  # nothing to do here, but remember this function for later!
```

```{r}
OurData %>%
  ggplot(aes(x = Depressed)) + # which column are the residuals?
  geom_bar()  # nothing to do here, but remember this function for later!
```

```{r}
OurData %>%
  ggplot(aes(x = Education)) + # which column are the residuals?
  geom_bar()  # nothing to do here, but remember this function for later!
```

```{r}
OurData %>%
  ggplot(aes(x = Work)) + # which column are the residuals?
  geom_bar()  # nothing to do here, but remember this function for later!
```

```{r}
OurData %>%
  ggplot(aes(x = HHIncome)) + # which column are the residuals?
  geom_bar()  # nothing to do here, but remember this function for later!
```




